\documentclass{article}

\linespread{1.1}
\usepackage[utf8]{inputenc}
\usepackage[left=1.5in,right=1.5in,bottom=1in]{geometry}
\setlength\parindent{0pt}
\setlength{\parskip}{1em}
\setcounter{secnumdepth}{0}
\usepackage{outlines}
\usepackage{graphicx}
\graphicspath{ {imgs} }
\usepackage{hyperref}
\usepackage{color,soul}

\newcommand{\entry}[3]{\textbf{#1}\quad #2. \\\textit{Keywords: #3}}

\title{Title
\\[5ex]Subtitle \\[3ex]}
\author{Carla Hyenne}
\date{January 16th, 2021}

\title{Real Life Examples of Urbanism}
\author{Carla Hyenne}

\begin{document}

\maketitle

\tableofcontents

\pagebreak

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Housing}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\entry{Leuven Beguinages}{jello world}{helwer}

\entry{Community Land Trusts}{hellwo world}{}

\entry{Billionaire's Row, NYC}{}{}

\entry{AirBnB}{}{rent gap theory, capitalism, spatial fix (Harvey)}

\subsection{Neighbourhoods}

\entry{Gated communities}{}{}

\entry{Suburbs}{American dream, for example in LA's suburbanisation. Belgian suburbs along highways, Belgians are `born with a brick in their stomach'}{}

\entry{Favelas, Rio}{}{}

\entry{Townships, South Africa}{}{}

\entry{Slums, Bangalore}{}{}

\entry{American ghetto}{}{}

\subsection{Housing Crisis}

\entry{Evergrande, China}{}{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Buildings and Districts}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Public Private Partnerships}

\entry{Tour et Taxis, Brussels}{}{}

\entry{Øresund bridge, Malmö/Copenhagen}{}{}

\subsection{Buildings}

\entry{Petrona Towers, KL}{}{}


\entry{North Quarters, Brussels}{}{Bruxellisation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Transport}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\entry{Fare Free Public Transport}{whelow owlrd}{}

\entry{HS2, UK}{https://www.bbc.com/news/uk-england-59333867}{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Lifestyle}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{New lifestyles}

\entry{The 15 minute city}{https://www.bloomberg.com/news/articles/2021-03-02/the-downsides-of-a-15-minute-city}{}

\subsection{Protest}

\entry{Detroit 1967}{ Detroit riots, confrontation of black residents and Detroit police}{critical urbanism}

\entry{Marolle 1969}{La Marolle protest, by residents against plans to expand the justice palace into the Marolle quartier}{critical urbanism}



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Culture}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\entry{Oktoberfest}{}{temporary space use}

\entry{Mardi Gras festivals}{}{ephemeral use of streets}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Societies and cities}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{}

\entry{(Post) socialist city}{Astana, Moscow}{}

\entry{Capitalist/Consumerist city}{LA, Vancouver}{}

\entry{Imperial cities}{Tokyo}{}

\entry{(Post) colonial}{}{}

\entry{(Post) industrial}{Roubaix boomed in the 20th century as a ville-cheminée (stacks), where work, home, leisure spaces were in the same place, but is now one of the poorest cities in France}{}

\entry{Mediterranean cities}{}{}

\entry{Well-managed cities}{Singapore (ref. Glaeser, 2011)}{}

\entry{Smart city}{Boston, Minneapolis, Milan (ref. Glaeser, 2011)}{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Urbanism concepts}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Spatial development}

\entry{Uneven spatial development}{The global North and South divide, historically reinforced by colonial power, today by post-colonial power.

Gentrification

Detroit}{gentrification, global North, global South, post-colonialism}

\entry{Privatised public space}{}{}

\subsection{Neoliberal city}

\entry{Creative destruction}{Penn Station, NYC. 

Place De Brouckère, Brussels: project to remove traffic from the boulevard, and undo the work from the 60s where axes of transit (automobile) were built in to the city, and people were de-prioritised. 

Senne, 19-20th century: Haussmann copy/paste urbanism approach was adopted in Brussels and the Senne was covered up.

Haussmann's Paris, 1850s-1860s: Haussmann develops profitable outlets like the Grands Magasins, the Opera. There is a massive transformation from the medieval character of Paris (narrow, dirty streets) to grand, clean avenues. This type of development requires a massive amount of capital. Engels questions the legitimacy of Haussmann: the plan was to make Paris more liveable, but did it consider the people who were currently living there? Will they be back in the new buildings that replaced the ‘slums’ and ‘ghettos’?

The City of London \href{https://journals.sagepub.com/doi/pdf/10.1068/b3401ed}{The creative destruction of cities}}{}

\entry{Extended urbanisation}{Alberta, the Texas of Canada supplying oil to the country and internationally.}{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Capitalism}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\entry{Geography of capitalism}{capitalism is embedded in varying social, cultural, institutional configurations. For example, colonial capitalism, war capitalism, state-controlled capitalism, capitalism with Chinese characteristics, etc.}{}

\entry{China's Evergrande}{he collapse of China’s second biggest property developer created fear that China’s financial system could collapse, however, it did not}{}


\if{False}

\entry{}{}{}

\fi

\end{document}
